---
title: "lab_3"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note to self: 

1    county|               county identifier
2      year|                            1987
3    crmrte|     crimes committed per person
4    prbarr|         'probability' of arrest
5   prbconv|     'probability' of conviction
6   prbpris| 'probability' of prison sentence
7    avgsen|             avg. sentence, days
8     polpc|               police per capita
9   density|             people per sq. mile
10    taxpc|          tax revenue per capita
11     west|           =1 if in western N.C.
12  central|           =1 if in central N.C.
13    urban|                   =1 if in SMSA
14 pctmin80|            perc. minority, 1980
15     wcon|       weekly wage, construction
16     wtuc|    wkly wge, trns, util, commun
17     wtrd| wkly wge, whlesle, retail trade
18     wfir|    wkly wge, fin, ins, real est
19     wser|      wkly wge, service industry
20     wmfg|         wkly wge, manufacturing
21     wfed|         wkly wge, fed employees
22     wsta|       wkly wge, state employees
23     wloc|        wkly wge, local gov emps
24      mix| offense mix: face-to-face/other
25  pctymle|              percent young male

```{r pressure, echo=FALSE}
library(tidyverse)
library(effsize)
library(plyr)
library(reshape2)

setwd("C:\\Users\\winbase\\MIDS\\w203\\lab_3-master")
crime_data = read.csv("crime_v2.csv")

#clean data, dropping all NA rows
crime_data = crime_data[which(crime_data$county != 'NA'),]

#very initial looking for correlations
cor(crime_data$crmrte, crime_data$taxpc) #0.45, pretty strongly correlated
cor(crime_data$crmrte, crime_data$wmfg) #0.35, also pretty strongly correlated
cor(crime_data$crmrte, crime_data$prbarr) #-0.39, pretty neg strong

# what is wrong with prbconv?
class(crime_data$prbconv[1][1]) #r thinks it's a factor?

```

We seem to have some encoding errors in the provided data set so we need to remedy that by making sure that all of the data is numeric.

```{r}

# let's clean these up
crime_data$prbconv = as.numeric(levels(crime_data$prbconv)[as.integer(crime_data$prbconv)])
 #not sure if /2 is correct but a lot of these variables are greater than 1
crime_data$prbconv = crime_data$prbconv/2

```

We probably need to start removing outliers so that we have a better insight into the trends and relations:

```{r}

remove_outliers_sd = function(d, field, numDevs) {
  s = numDevs * sd(d[[field]])
  m = mean(d[[field]])
  return( d[ which( d[[field]] > (m-s) & d[[field]] < (m+s)),])
}


#does crime rate go down when convictions are more likely?
#predict crmrte with conviction rate
outliers_removed_convictions = remove_outliers_sd(crime_data, 'prbconv', 2.0)
crime_conviction = lm(outliers_removed_convictions$crmrte ~ outliers_removed_convictions$prbconv)
plot(jitter(outliers_removed_convictions$prbconv), jitter(outliers_removed_convictions$crmrte), xlab = "Crime Rate", ylab = "Probability of conviction")
abline(crime_conviction) #doesn/t seem to show much

#predict crmrte with tax income
outliers_removed_tax = remove_outliers_sd(crime_data, 'taxpc', 2.0)
crime_tax = lm(outliers_removed_tax$crmrte ~ outliers_removed_tax$taxpc)
plot(jitter(outliers_removed_tax$taxpc), jitter(outliers_removed_tax$crmrte), xlab = "Crime Rate", ylab = "Tax receipts")
abline(crime_tax)
#summary(crime_tax)

#predict crmrte with police presence
outliers_removed_prbarr = remove_outliers_sd(crime_data, 'prbarr', 2.0)
crime_arrest = lm(outliers_removed_prbarr$prbarr ~ outliers_removed_prbarr$crmrte)
plot(jitter(outliers_removed_prbarr$crmrte), jitter(outliers_removed_prbarr$prbarr), xlab = "Crime Rate", ylab = "Probability of Arrest")
abline(crime_arrest)

#maybe remove the outliers for polpc?
outliers_removed_polpc = remove_outliers_sd(crime_data, 'polpc', 2.0)
crime_police = lm(outliers_removed_polpc$polpc ~ outliers_removed_polpc$crmrte, data = outliers_removed_polpc)
plot(jitter(outliers_removed_polpc$crmrte), jitter(outliers_removed_polpc$polpc), xlab = "Crime Rate", ylab = "Police Presence")
abline(crime_police)

```

We can see a as police presence increases, so does our crime rate. Police presence could be required by an increased crime rate of course, but we can also see some evidence that it does not seem to reduce the crime rate significantly. There is evidence that probability of arrest seems to correspond to a lower crime rate. 

```{r}



```

Note that the `echo = FALSE` parameter was added to the code chunk to prevThisent printing of the R code that generated the plot.

```Test code from Ryan```
